---
title: 'HUDK4050: Class Activity 6'
author: "Ningyao Xu"
date: "10/16/2018"
output: pdf_document
---
# Data Management
```{r}
#Load data
DF1 <- read.csv("HUDK405019-clustering.csv", header = TRUE)
library(tidyr)
library(dplyr)
#rownames == First name + Last name
for (i in c(1,2,15,16))
{DF1[,i] = as.character(DF1[,i])}
DF1$name <- paste(DF1$First.Name, DF1$Last.Name)
rownames(DF1) <- DF1$name
DF1 <- DF1[,3:16]


#Delete those who write latitude and longtitude twice in the survey
list <- NULL
for (i in 1:nrow(DF1))
  { if (DF1[i,13] == DF1[i,14] )
list <- c(list,i)}
DF1 <- DF1[-list,]


# reverse those who put latitude and longtitude in wrong order
a <-NULL
b<- NULL
reverse <- grep("E",DF1[,13])
for (i in reverse)
  { a =  DF1[i,13]
    b =  DF1[i,14]
    DF1[i,13] <- b
    DF1[i,14] <- a}

#Find the signal and delete all the things after the signal
#"Â° is how my DELL shows "°", I have no idea why it shows this way
for (j in c(13:14)){
for (i in 1:nrow(DF1))
  { if (grepl("Â°", DF1[i,j]) )
  { psn <- as.numeric(regexpr("Â°", DF1[i,j]))
    DF1[i,j] <-  substr(DF1[i,j], 1, psn-1)}}}

#If you are using mac, use the following one
for (j in c(13:14)){
for (i in 1:nrow(DF1))
  { if (grepl("\\D", DF1[i,j]) )
  { psn <- as.numeric(regexpr("\\D", DF1[i,j]))
    DF1[i,j] <-  substr(DF1[i,j], 1, psn-1)}}}

#Delete all the space, alphabet from the data and turn all the data into numeric 
 for (i in c(1:11,13,14))
{ DF1[,i]= gsub("[[:alpha:]]", "", DF1[,i]) 
  DF1[,i]= gsub(" ", "", DF1[,i])
  DF1[,i] = as.numeric(DF1[,i])}
# Omit all the NAs from the data
DF1 <- na.omit(DF1)

DF2 <- data.frame(select_if(DF1,is.numeric))
#Convert the index numbers of the data fram into the student names.
  
#Wrangle data using dplyr to include only the numerical values.

#Scale the data so that no variable has undue influence

DF2 <- scale(DF2)

```

# Find lattitudes & longitudes for cities
```{r}
#Unfortunately Google has restricted access to the Googple Maps API so the code below no longer works. Instead you have the lats and longs in your data.

#install.packages("ggmap")
#install.packages("rgdal")
#library(ggmap)
#library(tmaptools)

#Request lattitude and longitude from Google Maps API
#DF2 <- geocode(as.character(DF2$Q1_1), output = "latlon", source = "dsk")
```

Now we will run the K-means clustering algorithm we talked about in class. 
1) The algorithm starts by randomly choosing some starting values 
2) Associates all observations near to those values with them
3) Calculates the mean of those clusters of values
4) Selects the observation closest to the mean of the cluster
5) Re-associates all observations closest to this observation
6) Continues this process until the clusters are no longer changing

Notice that in this case we have 10 variables and in class we only had 2. It is impossible to vizualise this process with 10 variables.

Also, we need to choose the number of clusters we think are in the data. We will start with 4.

```{r}

fit <- kmeans(DF2, 4) 

#We have created an object called "fit" that contains all the details of our clustering including which observations belong to each cluster.

#We can access the list of clusters by typing "fit$cluster", the top row corresponds to the original order the rows were in. Notice we have deleted some rows.

fit$cluster

#We can also attach these clusters to te original dataframe by using the "data.frame" command to create a new data frame called K4.

DF3 <- data.frame(DF2, fit$cluster)

#Have a look at the DF3 dataframe. Lets change the names of the variables to make it more convenient with the names() command.


```

# Visualize your clusters in ggplot
```{r}
#Create a scatterplot that plots location of each student and colors the points according to their cluster 
DF4 <- data.frame(DF1[,13],DF1[,14],fit$cluster)
names(DF4)  <- c("latitude", "longtitude","cluster")
attach(DF4)
library(ggplot2)
ggplot(DF4, aes(x =  longtitude, y =latitude, pch = factor(cluster))) +
    geom_point(aes(color = factor(cluster)))

```

# Can you group students from the classes data set in Assignment 2 using K-modes?

```{r}

DT1 <- read.csv("hudk4050-classes.csv",header = TRUE)
DT1$Name <- paste(DT1$First.Name, DT1$Last.Name)
DT2_dirty <- DT1[,3:9]
DT3 <- DT2_dirty %>% gather(classnum, classcode, `Class.1`, `Class.2`, `Class.3`, `Class.4`, `Class.5`, `Class.6`) %>% select(-c(classnum))
DT3$classcode = gsub(" ", "", DT3$classcode)
DT3 <- DT3 %>% filter(classcode != "HUDK4050") %>% filter(Name != "ZIMO CHEN")
DT3$Count = 1
DT3 <- DT3[which(DT3$classcode != ""),]
DT4 <- DT3 %>% spread(classcode,Count)
row.names(DT4) = DT4$Name
DT4$Name <- NULL
DT4 = ifelse(is.na(DT4), 0, 1)
DT5 = as.matrix(DT4)
DT5 <- scale(DT5)
library(MASS)
set.seed(123)
pca=princomp(DT5[,1:50],cor=T)
screeplot(pca,type="line",lwd=2)
#According to this plot, maybe we should try cluster 6 groups.
set.seed(123)
fit2 <- kmeans(DT5,6) 
fit2$size
cluster <- data.frame(fit2$cluster)
colnames(cluster) <- c("cluster")
cluster

# Just to check, I am in the Applied Statistics program and people in the group 5 are exactly those in the Applied statistics program with me!

```

